---
title: "Machine Learning Project"
author: "R.L. Rojas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(caret)
library(dplyr)
library(arsenal)
# Loading in Data

training = read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")

testing = read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")

```



# Cleaning Up Data
Some of the variables in the testing and training don't give us any information, so we're going to remove them. 
```{r warning=FALSE, message=FALSE}

vars.to.remove = c("X", "user_name", "new_window", "num_window", "cvtd_timestamp", "problem_id", "raw_timestamp_part1", 
"raw_timestamp_part2")


training = training[,names(training) %nin% vars.to.remove]

testing = testing[,names(testing) %nin% vars.to.remove]

```


Now converting missing data to zeros (can't having missing data with Random Forests)

```{r warning=FALSE, message=FALSE}

training = training %>% mutate_all(~ ifelse(is.na(.) | . == "" | . == " " | . == '#DIV/0!', 0, .))


testing = testing %>% mutate_all(~ ifelse(is.na(.) | . == "" | . == " " | . == '#DIV/0!', 0, .))


training$classe = factor(training$classe, levels = c(1, 2, 3, 4, 5), 
    labels = c("A", "B", "C", "D", "E"))

```


# Setting Seed
```{r}
set.seed(444)
```


# Running Tree (CART) Model
```{r warning=FALSE, message=FALSE}
mod1 = train(classe ~ ., data = training, method = "rpart")

confusionMatrix(mod1)
```


The accuracy for the first model using just a tree is about 50%, meaning the out of sample error is also about 50%. Not really an accurate model...

Bagging should help get a more accurate model, so the method "treebag" will be used. 

# Running Bagged Tree
```{r, warning=FALSE, message=FALSE}
mod2 = train(classe ~ ., data = training, method = "treebag")

confusionMatrix(mod2)
```
 

The accuracy of this model is much, much higher than the non-bagged CART with a .59% out of sample error. 

# Use Simple Tree Model to Predict Exercise 
```{r warning=FALSE, message=FALSE}
predict(mod1, testing)
```


# Use Bagged Tree Model to Predict Exercise
```{r warning = FALSE, message = FALSE}
predict(mod2, testing)

```